---
title: "MathAssignment4"
author: "Christina"
date: "July 18, 2015"
output: html_document
---

1. The system of equations can be expressed as: \
a: 
```{r echo=FALSE}
matrix(c(1,2,-1,2,1,1,-3,-3,0), nrow=3)
```
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;b:
```{r echo=FALSE}
matrix(c(5,13,8))
```
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;x:
```{r echo=FALSE}
matrix(c('x','y','z')) 
```
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;The solutions for x,y,z can be found by solving ax=b \
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;Matrix a cannot be inverted. In fact, it concains overlapping information: row 1&2 vs. row 3. The system is exactly singular.  

2. The following function sol attempts to provide a solution for the above problem, ax=b.
```{r}
sol <- function(a,b) { 
  if (ncol(a) != nrow(b)){
    print ("Error: check input matrix dimensions")
  }
  else {
    test <- tryCatch(solve(a,b), error = function(e) e) 
      if (any(class(test) == "error")){
        print ("Something went wrong: For example, a may not be inversible")
      }
      else {
        solve (a,b)
    }
  }
}
```
&nbsp;&nbsp;&nbsp;&nbsp;Test code:
```{r}
a <- matrix(c(1,2,-1,2,1,1,-3,-3,0), nrow=3)
b <- as.matrix(c(5,13,8))
sol(a,b)
```
3. Manual calculation:  
``AB[11]=4*1-3*3=-5
    AB[12]=4*4+3*2=22 
    AB[21]=-3*1+5*3=12
    AB[22]=-3*4-5*2=-22 
    AB[31]=3
    AB[32]=-2``

```{r echo=FALSE}
matrix(c(-5,12,3,22,-22,-2), nrow=3)
```

4. Using R to solve AB:
```{r}
A <- matrix(c(4,-3,0,-3,5,1), nrow=3)
B <- matrix(c(1,3,4,-2), nrow=2)
A %*% B
```
