---
title: "RAssignment2"
author: "Christina Taylor"
date: "July 8, 2015"
output: html_document
---

### Function "fact": Calculate Factoral

**Factoral is defined as n!. The user defined function "fact" accomplishes this calculation. It takes an integer argument x. If x is not an integer, an error message is given.**  

See the code to build the function here: 
```{r}
fact<-function(x)
{
i<-1
result<-1

	if (is.numeric(x)==FALSE) 
	{
	  print ("you did not enter a number")
	}

	else if (x==0) 
	{
	  return (1)
	}
	
	else if (x%%1==0 && x>0)
	{
		for (i in 1:x)
		{
		  result<-result*i
		  i<-i+1
		}
		return (result)
	}
	
	else 
	{
	  print ("you did not enter an integer")
	}
}
```

See tests and results here:
```{r}
fact(0)
fact(5)
fact(5.5)
fact("factoral")
```

### Function "choo": Evaluate "n choose r"
**In probability, "n choose r" evaluates to n! / (( n - r)! * r!). The user defined function "choo" accomplishes this calculation. It takes 2 integer arguments, n and r, and return the evaluation. If n or r is not an integer, an error message is given.  If n is less than r, the fucntion evaluates to 0.**

See the code to build the function here: 
```{r}
choo<-function(n,r)
{
  if (is.numeric(n)==FALSE || is.numeric(r)==FALSE) 
  {
    print ("you did not enter numbers")
  }
    
  else if (n%%1==0 && r%%1==0)
	{
		if (n >= r)
		{
			eval <- fact(n) / fact(r) / fact(n-r) 
		}
		else
		{
			eval <- 0
		}	
		return (eval)
	}
	
  else
	{
	  print ("you did not enter integers")
	}
}
```

See tests and results here:
```{r}
choo(5,3)
choo(3,5)
choo(3.5,5)
choo("choose")
```
