---
title: "RAssignment4"
author: "Christina"
date: "July 18, 2015"
output: html_document
---

Mtcars Dataset: It was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 other aspects of 32 automobiles' performance.

```{r}
#load data
library(datasets)
data(mtcars)
```
Dataset at a glance:
```{r}
#identify variable type
str(mtcars)
#generate summary level description
summary(mtcars)
```

Histrogram of mileage per gallon distribution:
The graph shows the frequency distribution of fuel efficiency.
```{r, message=FALSE, echo=FALSE}
require(ggplot2)
require(scales)
require(ggthemes)
g1<-ggplot(mtcars) + geom_histogram(aes(x=mpg),fill="grey50")
g1<-g1+labs(title="fuel efficiency distribution", x="miles/gallon", y="number of cars")
g1+theme_tufte()
```

Boxplot of gross horsepower distribution by number of cylinders:\
The graph shows the variance of engine performance. More cylinders suggests more powerful engines. 
```{r, message=FALSE, echo=FALSE}
g2<-ggplot(mtcars, aes(x=as.factor(cyl),y=hp))+geom_boxplot()
g2<-g2+labs(title="engine performance distribution", x="number of cylinders", y="gross horse power")
g2+theme_tufte()
```
\
Scatterplot of horsepower and mileage per gallon colored by number of cylinders:\
The graph shows a negative correlation between engine performance and fuel consumption. 
```{r, message=FALSE, echo=FALSE}
g3<-ggplot(mtcars, aes(x=mpg,y=hp))
g3<-g3+geom_point(aes(color=cyl))
g3<-g3+labs(title="engine performance vs. fuel efficiency", x="miles/gallon", y="gross horse power", color="number of cylinders")
g3+theme_tufte()
```
