---
title: "Hflights Data Analysis"
author: "Christina"
date: "July 19, 2015"
output: html_document
---

**I want to choose the travel time and airline smartly, so I have the best chance to avoid delays and arrive at my destination in time. So I analyzed the hflight data with the KPI Arrival Delay:**  

```{r message=FALSE}
#load data
library("hflights")

#change column data type
hflights$UniqueCarrier<-as.factor(hflights$UniqueCarrier)
hflights$DayOfWeek<-as.factor(hflights$DayOfWeek)

#create and append departure time field
FlightDate<-paste(hflights$Year, hflights$Month, hflights$DayofMonth,sep="-")
data<-cbind(hflights,FlightDate)

#select rows where flights are delayed
del<-data[which(data$ArrDelay >0),-(1:3)]

#subset columns of interest
sub<-subset(del, select=c(DayOfWeek,UniqueCarrier,ArrDelay,DepDelay,AirTime,FlightDate))

#summary statistics for data analysis
summary(sub)
```

**1. Histogram of Arrival Delays by day of week:**  

```{r, message=FALSE}
#qplot(UniqueCarrier, data=sub, geom="bar", fill=DayOfWeek)
#ggplot(sub, aes(UniqueCarrier)) + geom_freqpoly(aes(group = DayOfWeek, colour = DayOfWeek))
#Stacked bars are easy, but might be overloaded with information. Faceting might be a better solution.
```
```{r, message=FALSE, echo=FALSE}
require(ggplot2)
require(ggthemes)
require(scales)
g1<-ggplot(sub, aes(DayOfWeek, fill="red")) + geom_bar() + facet_wrap(~ UniqueCarrier)
g1<-g1+labs(title="Flight Delays by Day/Airline", x="Day of Week", y="Number of Delayed Flights") 
g1+theme_gdocs()+guides(fill=FALSE)
```
\
\
Congestion tends to happen during weekend and mid week, although there are viarations for airlines with smaller volume.\
\
\
**2. Boxplot of Delay Time by airline:**  
```{r, message=FALSE, echo=FALSE}
g2<-ggplot(sub, aes(x=UniqueCarrier,y=ArrDelay))+geom_boxplot()
g2<-g2+labs(title="Delay Time by Airline", x="Airline", y="Delay Time (min)")
g2+theme_gdocs()
```
\
\
Most of the delays are short (mean = 24 min), although there are significant outliers with certain airlines. \
\
\
**3. Evaluate airline performance:**   
Simply counting the frequency of delayed flights per airline is misleading, as bigger outlines represent majority of flights and likely most of the delays. Therefore, I created calculated fields for analysis:   
```{r, message=FALSE}
#calculate number of delayed flights by carrier
library("sqldf")
ucd<-sqldf("select UniqueCarrier, count(ArrDelay) as delays from hflights where ArrDelay >0 group by UniqueCarrier")

#calculate number of total flights by carrier, if arrival information exists
uc<-sqldf("select UniqueCarrier, count(ArrDelay) as flights from hflights where ArrDelay is not null group by UniqueCarrier")

#merge the results and create calculated field: %of delayed flights
library(dplyr)
stats<-merge(ucd,uc,all=TRUE)
ratio<-with(stats, 100*delays/flights)
stats<-cbind(stats,ratio)
stats
```
\
Bar chart of percentage of delayed flights, organized by airline size; number of flights labeled: \
```{r, message=FALSE, echo=FALSE}
g3<-ggplot(stats, aes(x=(reorder(UniqueCarrier,-flights)),y=ratio, fill = ratio)) + scale_fill_gradient2(high=muted("red")) + geom_bar(stat = "identity") + geom_text(size=3, data=stats, aes(label = flights), hjust=0) + coord_flip() 
g3<-g3+labs(title="Flight Delays", x="Airline (smallest top)", y="Delay Rate", fill="Delay Rate")
g3+theme_gdocs()
```
\
\
AA has the lowest arrival delay rate while F9 the highest. Notice that from a certain point on, flight delay rate tend to increase as the volume of flights grows. This can also be seen in the following scatter plot:
\
```{r, message=FALSE, echo=FALSE}
g4<-ggplot(stats, aes(x=flights,y=ratio))+geom_point(aes(color=UniqueCarrier))+geom_text(size=3, data=stats, aes(label = UniqueCarrier), hjust=0)
g4<-g4+labs(title="Flights Volume vs. Delay Rate", x="Airline Size", y="Delay Rate")
g4+theme_gdocs()+guides(color=FALSE)
```

\
\
**4. (Is it obvious?) Scatter plot of arrival delay and departure delay colored by airline:**    
```{r, message=FALSE, echo=FALSE}
g5<-ggplot(sub, aes(x=DepDelay,y=ArrDelay))
g5<-g5+geom_point(aes(color=UniqueCarrier))
g5<-g5+labs(title="Arrival vs. Departure Delay", x="Departure Delay (min)", y="Arrival Delay (min)", color="Airline")
g5+theme_gdocs()
```
\
There is strong positive correlation between departure delays and arrival delays, suggesting airport congestion rather than flight time likely caused the delay.  
\
**My conclusion is that I want to avoid weekends; Tuesday/Thursday may be best. I would also choose to fly US, FL, or AA - they have decent number of flights but low delay rate.**
